### 2018-04-13 ident/docs/draft_2.tex
completed section 3.2

	\textcolor{red}{summarize and start next section on identifiability here}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$). Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) from the network shown in Figure \ref{fig:network}.

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}	
	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition. This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC vs Convenience kinetics) to generate in silico experimental data. We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (Figure \ref{fig:ident}a). Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters. In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (Figures \ref{fig:ident}b and f).

	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter. To elaborate, we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them. So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.

	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data. However, the difference in data usefulness between similar experiments due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models. The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets and its subsequent effect of parameter identifiability. Nonetheless, we would like to note that these differences in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments. We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}). Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}) when enzyme concentrations are available.

@@

	 In contrast, when $v_3$ is modeled using the Convenience Kinetic rate law, we find that the parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}. We say locally identifiable due to the fact that all three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two possible roots. In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$. The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.

	The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d. The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in Figure \ref{fig:ident_root2}c.

	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ (Figure \ref{fig:ident_root2}b) are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$. We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}). We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c). In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d). Thus, we can get globally identifiable parameters through the addition of constraints related to their physical relevance to locally identifiable parameters.

	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters. This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b). Instead, $K_3^{fdp}$ becomes globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}a), and $K_3^{pep}$ becomes globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}b). Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable. Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability. This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.

	Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.

@@

	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).

	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters. This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b). Instead,

	 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable. Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability. This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.

	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model. For instance, as we state in the Methods section,

	 We say locally identifiable due to the fact that

	 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$. The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.

	 Accordingly, we can estimate unique values for any globally identifiable parameter from any set of experiments that satisfy practical identifiability criteria.

@@

	For flux $v_1$ (Figure \ref{fig:data_utility}a), data from about 52\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$). Only $V_1^{max}$ can be identified by about 7\% of the combinations (Figure \ref{fig:ident}a), while only about 40\% of the combinations can identify both the parameters used in the model. 	
	In contrast, all the available data can identify both parameters in $v_2$ (Figure \ref{fig:data_utility}b). This is in agreement with the degree of identifiability of parameters used to model $v_2$ (Figure \ref{fig:ident}b). In the case of $v_3$ (Figure \ref{fig:data_utility}c), close to 60\% of the \textcolor{red}{4896 combinations} can identify only one parameter, while less than 10\% of the combinations can identify all three parameters in the model. Note that only 5\% of combinations cannot identify any parameter modeling $v_3$.

	Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation. For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative. This can lead to potential waste of resources used to perform these experiments.

	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} also informs the ease of identifying parameters for a single flux. For the small network (Figure \ref{fig:network}), we find that all chosen combination of experiments are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:data_utility}b). In contrast, only about 40\% of the combinations can identify both parameters in $v_1$. Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ and $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$. A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$. A similar argument can also be for determining experiments that can identify $V_e^{max}$ and $K_e^{fdp}$ in $v_5$: since $V_e^{max}$ has an identifiability of 100\%, the choice of experiments to completely identify parameters in $v_5$ rests with experiments that can identify $K_e^{fdp}$.

	We believe that this information (Figure \ref{fig:data_utility}), in conjunction with the identifiability of each parameter presented in Figure \ref{fig:ident} will give an idea on the hierarchy of identifiability of each parameter, and the need for experimental design to identify each parameter. Additionally, having the ability to see the contribution of each type of experiment towards identifiability (Figure \ref{fig:exp_info}) would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism. So, we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network in the following section.

	%Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information. Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.

	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}		
	From Figure \ref{fig:ident} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model. For example, the maximum reactions rates of $v_2$ and $v_5$ have a degree of identifiability of 100\% (Figure \ref{fig:ident}b and d), while their counterparts $V_1^{max}$ and $V_3^{max}$ in $v_1$ and $v_3$ respectively (Figure \ref{fig:ident}a and c), are only identifiable to a lesser extent.	
	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified. In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}. Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range. %Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals. In linear systems, persistence of excitation of any input signal can be theoretically guaranteed. However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis. The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.

@@

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_data_utility}}
		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters. Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$. The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments. The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux. A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, \textcolor{red}{3360 combinations} are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$. Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
	\end{figure}

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_v5_1_data_utility}}
		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters. Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$ and c) $v_5$. The total number of combinations of experiments is shown in the vertical axis, and the horizontal axis shows the maximum number of parameters that be identified using data from a combination of a) two, b) three and c) two experiments. The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux. A total of 240 data combinations are used for identifiability analysis for a) $v_1$, \textcolor{red}{3360 combinations} are used to practically identify b) $v_3$, and 420 combinations are used to analyze the identifiability of c) $v_5$. Information processed based on the identifiability of a) $v_1$ in the presence of enzyme concentration data, b) the second root of parameters in $v_3$ and c) the first root of $K_e^{fdp}$ in $v_5$ are shown.}\label{fig:other_data_utility}
	\end{figure}

>>>>

	So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$). Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) based on the expressions and the identifiable roots we discussed in this section.

	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}		
	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}. Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter. We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models. We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition. This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) to generate in silico experimental data. We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.

	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (47.5\% and 40.0\% in Figure \ref{fig:ident}a). Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (25\% in Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters. In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (100\% in Figures \ref{fig:ident}b and f).

	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter. To elaborate, we showed in Equation (\ref{eq:v1_par}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them. So, in this instance, recall that the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}. We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section x.

	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data. However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models. The interaction between different components of the network (Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network. The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.

	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments. We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}). Recall that any predicted parameter value for both $k_1^{cat}$ and $V_1^{max}$ matches with the exact known value irrespective of the experiment (MWC vs Convenience kinetics) from which data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).  Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.

@@

	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figure \ref{fig:ident}c and g). Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources. In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable (Section \ref{sec:example}: $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$, $K_3^{pep}$are globally identifiable when $K_3^{fdp}$ is known and fixed) in Supplementary Figure \ref{fig:ident_v3_var}. Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model results in only a fraction of data sets being useful to estimate parameters (Supplementary Figure \ref{fig:ident_v3_var}a and b). This ties back with our earlier observation that these parameters are not practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b). However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model. This can be corroborated by the parameter values shown in Supplementary Figure \ref{fig:parameter_value_v3_var}c and d, where data from the Convenience kinetics model is used for analysis.

	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h). The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}c and f.

@@

@@

	For $v_1$ (Figure \ref{fig:data_utility}a), data from about 50\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$). In contrast, data from only about 5\% of the experiments is useless towards identifying any parameter in $v_3$, while there are no useless combinations of experiments from which data cannot be used to identify either $v_2$, or $v_5$. These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}). Thus, when numerous experiments are possible to obtain steady state data to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.

	We see that in instances where the identifiability of the fluxes differ based on the data set used (data from MWC model vs data from Convenience kinetics model), the utility of the data sets towards identifying parameters are changes. For instance, when data from the MWC model is used to identify parameters in $v_3$ (Figure \ref{fig:data_utility}c), less than 10\% of the data sets can identify all three parameters. Whereas, when data from the Convenience kinetics model is used for the same purpose, slightly more than 50\% of the combinations can identify all three parameters. Note that this observations should be considered in light of the structural local identifiability and the practical non-identifiability that we presented earlier in Sections \ref{sec:example} and \ref{sec:ident_rate_laws}.

	Although one could argue based on earlier observations that uncertainty in the model from which data is obtained, and the structural non-identifiability of some of the parameters in the models of the aforementioned fluxes that there may not be a need for experimental design if in vivo model uncertainty and structural non-identifiability can be resolved, one needs to only look at Supplementary Figure \ref{fig:other_data_utility}a and d. In these figures we show the utility of multiple combinations of metabolite and protein concentration, and flux data towards identifying $k_1^{cat}$ and $K_1^{ac}$ in $v_1$ when data from the MWC model and Convenience kinetics model are used, respectively. We see that about 45\% of the 240 combinations of data from 21 different experiments cannot identify either one of the two parameters.

	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} and Supplementary Figure \ref{fig:other_data_utility} also informs the ease of identifying parameters for a single flux. In conjunction with the degrees of identifiability shown in Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we see that identifying $v_2$ is much easier than identifying $v_1$. Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ and $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$. A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$. A similar argument can also be for determining experiments that can identify $V_e^{max}$ and $K_e^{fdp}$ in $v_5$: since $V_e^{max}$ has an identifiability of 100\%, the choice of experiments to completely identify parameters in $v_5$ rests with experiments that can identify $K_e^{fdp}$.

	So, if we can establish a hierarchy of identifiable parameters and the corresponding fluxes they model, this section gives credence to the idea that a priori experimental design is necessary, even when all possible in vivo measurements can be obtained, to get the most informative data for parameter identification and estimation at minimum cost. Additionally, having the ability to see the contribution of each type of experiment towards identifiability would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism. So, we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network in the following section.

	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}		
	From Figures \ref{fig:ident} and \ref{sec:data_utility} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model. For example, the maximum reactions rates of $v_2$ and $v_5$ have a degree of identifiability of 100\% (Figure \ref{fig:ident}b and d), while their counterparts $V_1^{max}$ and $V_3^{max}$ in $v_1$ and $v_3$ respectively (Figure \ref{fig:ident}a and c), are only identifiable to a lesser extent.	
	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified. In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}. Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range. %Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals. In linear systems, persistence of excitation of any input signal can be theoretically guaranteed. However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis. The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.

@@

	%Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation. For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative. This can lead to potential waste of resources used to perform these experiments.

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters. Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$. The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments. The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux. A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$. Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
	\end{figure}

@@
